### 2022-04-18 GB_Simulation_3Species.Rmd
adult weight 1

---
  title: "Modeling 3 species on George's Bank (Yellowtail Flounder, Cod, Haddock)"
editor_options: 
  chunk_output_type: console
---

```{r}
library(Rcpp)

#just sum matrix for denominator and then do x1 / sum_mat. 1 year took 4.059 minutes
cppFunction('

            double sum_mat(NumericVector M){

            double S = sum(M);

            return S;
            }
            ')

sum(hab$hab$spp1[nonzero_idx[[1]]])

sum_mat(hab$hab$spp1[nonzero_idx[[1]]])

#normalize entire matrix in C++. 4.95 minutes for 1 year.
#I think because I must use as.matrix to entier MM. Will change to vector instead below

cppFunction('

NumericMatrix norm_mat(NumericMatrix M, NumericMatrix MM, NumericMatrix Nzero_vals) {
  //NumericMatrix MM = M(! is_na(M));
  double sm = sum(MM);
  int nNonzero = Nzero_vals.nrow();

  int nrow = M.nrow();
  int ncol = M.ncol();

  NumericMatrix new_val(nrow, ncol);// for storing new values

  for(int p = 0; p < nNonzero; p++)  {

    new_val(Nzero_vals(p,0)-1,Nzero_vals(p,1)-1) = M(Nzero_vals(p,0)-1,Nzero_vals(p,1)-1) / sm ;

  }

  return new_val;
}
')

#normalize entire matrix in C++. Send MM as vector instead of using as.matrix()
#took 4.71 minutes to run 1 year
#norm_mat(x1, x1[nz[[s]]],Nzero_vals = nz[[s]])
cppFunction('

NumericMatrix norm_mat(NumericMatrix M, NumericVector MM, NumericMatrix Nzero_vals) {

  double sm = sum(MM);
  int nNonzero = Nzero_vals.nrow();

  int nrow = M.nrow();
  int ncol = M.ncol();

  NumericMatrix new_val(nrow, ncol);// for storing new values

  for(int p = 0; p < nNonzero; p++)  {

    new_val(Nzero_vals(p,0)-1,Nzero_vals(p,1)-1) = M(Nzero_vals(p,0)-1,Nzero_vals(p,1)-1) / sm ;

  }

  return new_val;
}
')

#integrated the normalization into the movement command. This runs the fastest: 3.82 minutes for 1 year

```

  ## Habitat setup

First read in habitat created using point data and covariates (sediment and depth)
and create other required info

It returns a list of suitable habitat for each species (hab), the habitat as
adjusted during the spawning period (spwn_hab) and the binary location of
spawning areas (spwn_loc). It also returns the locations as x1,x2,y1,y2 and the
multiplier of attractiveness to the spawning area during spawning periods
(spwn_mult).

If plot.dist = TRUE, it returns the plots to a file.

```{r habitat}

#LOAD HABITAT PREVIOUSLY CREATED
#hab <- readRDS(file="hab_justYT.RDS") #finer resolution
#hab <- readRDS(file="hab_justYT2.RDS") #courser resolution

hab <- readRDS(file="hab_GB_3species.RDS") #courser resolution

#plot strata
par(mar=c(1,1,1,1))
fields::image.plot(hab$stratas)

#plot habitat
par(mar=c(1,1,1,1))
fields::image.plot(hab[["hab"]][["spp1"]])

par(mar=c(1,1,1,1))
fields::image.plot(hab[["hab"]][["spp2"]])

par(mar=c(1,1,1,1))
fields::image.plot(hab[["hab"]][["spp3"]])

#plot spawn habitat
par(mar=c(1,1,1,1))
fields::image.plot(hab[["spwn_hab"]][["spp1"]])

par(mar=c(1,1,1,1))
fields::image.plot(hab[["spwn_hab"]][["spp2"]])

par(mar=c(1,1,1,1))
fields::image.plot(hab[["spwn_hab"]][["spp3"]])

```

# Initialise the simulation

This vignette is a paired down example of how to construct a simulation using
MixFishSim. We include only a basic example and encourage users to explore the
other features of the package. \\

## Base parameters

First we specify the basic parameters of the simulation. This includes the
dimensions of the spatial domain, the number of years to simulate, the number
of fleets and vessels per fleet and the number of species and how often (in
                                                                         weeks) the fish move.

The object returned is used internally by MixFishSim a list with two levels:

  * sim$idx : The different units of different processes 
* sim$brk.idx: breaks for each of the key processes in units of a timestep

```{r basic}

#NEW VERSION that has week breaks for entire simulation
# source("R/init_sim_Bens.R")
# sim <- init_sim_Bens(nrows = 100, ncols = 100, n_years = 20, n_tows_day = 1,
#                 n_days_wk_fished = 1, n_fleets = 1, n_vessels = 0, n_species = 2,
#                 move_freq = 1)

#NEW VERSION that has week breaks for entire simulation and allows fishing on just 1 day per week
source("R/init_sim_Bens_nofish.R")
sim <- init_sim_Bens_nofish(nrows = nrow(hab$hab$spp1), ncols = ncol(hab$hab$spp1), n_years = 22, n_tows_day = 1,n_days_wk_fished = 1, n_fleets = 1, n_vessels = 1, n_species = 3,
                            move_freq = 1)

class(sim)
sim$idx
names(sim$brk.idx)

```

## Population models

Now we need to set up the population models for the simulations. We do this
with the init_pop function. We set the initial population biomasses, movement
rates, recruitment parameter and growth and natural mortality rates.

The object created stores all the starting conditions and containers for
recording the changes in the populations during the simulations.

We can plot the starting distributions for each population as a check.

```{r pop_init}

#hab <- readRDS("hab_GB_3species.RDS")

#rebuild package
devtools::build()

#load rcpp exports
Rcpp::sourceCpp(file= "src/Movement.cpp")
Rcpp::sourceCpp(file= "src/RcppExports.cpp")
Rcpp::compileAttributes() #this updates RcppExports.R file, which contains function definitions

#CALULATE INDICES OF NONZERO VALUES IN HAB TO PASS TO MOVE_POPULAITON DURING MOVEMENT
nonzero_idx <- lapply(paste0("spp", seq_len(sim$idx[["n.spp"]])), function(s) {

  which(hab[["hab"]][[s]] >0 , arr.ind=T)

})

names(nonzero_idx) <- paste("spp",seq_len(sim$idx[["n.spp"]]), sep ="")

 source("R/init_pop_Bens.R")
# profvis({

##where should starting point be? check highest values locations
# which(hab[["hab"]][["spp1"]]==max(hab[["hab"]][["spp1"]],na.rm=T),arr.ind=T)
# which(hab[["hab"]][["spp2"]]==max(hab[["hab"]][["spp2"]],na.rm=T),arr.ind=T)
# which(hab[["hab"]][["spp3"]]==max(hab[["hab"]][["spp3"]],na.rm=T),arr.ind=T)
# 
# #make sure each have nonzero value on start cell
# hab[["hab"]][["spp1"]][46,112]
# hab[["hab"]][["spp2"]][30,134]
# hab[["hab"]][["spp3"]][24,91]

#YELLOWTAIL, COD, HADDOCK

#profvis::profvis({
#decreasing population settings
Pop <- init_pop_Bens(sim_init = sim, Bio = c("spp1" = 3194, "spp2" = (13000+30000)/2, "spp3" = 150000*1.2), #CHANGE
                     hab = hab[["hab"]], start_cell = list("spp1" = c(46,112),"spp2" = c(30,134),"spp3" = c(24,91)),
                     lambda = c("spp1" = 0, "spp2" = 0, "spp3" = 0), #set at 0 first, change after initial population set
                     init_move_steps = 20,
                     rec_params = list("spp1" = c("model" = "BH", "a" = 30445, "b" = 4301, "cv" = 0),
                                       "spp2" = c("model" = "BH", "a" = 27868 , "b" = 3491, "cv" = 0),
                                       "spp3" = c("model" = "BH", "a" = 73568, "b" = 40530, "cv" = 0)), #CHANGE
                     rec_wk = list("spp1" = 9:12, "spp2" = 8:13, "spp3" = 11:14),
                     spwn_wk = list("spp1" = 9:12, "spp2" = 8:13, "spp3" = 11:14  ), 
                     M = c("spp1" = .2064+.358, "spp2" = .2728+.511, "spp3" = .334+.45), #NEED F
                     K = c("spp1" = .2295, "spp2" = .16, "spp3" = .2465),
                     Weight_PreRecruit = (c("spp1" = .13/.39, "spp2" = .39/2.95, "spp3" = .19/1.12)),
                     Weight_Adult = (c("spp1" = 1, "spp2" = 1, "spp3" = 1)),
                     nz = nonzero_idx)

#})

#})
# actual weights
#          Weight_PreRecruit = (c("spp1" = .13, "spp2" = .39, "spp3" = .19)),
 #                    Weight_Adult = (c("spp1" = .39, "spp2" = 2.95, "spp3" = 1.12)),

#set lambda at 0 first, change after initial population set
Pop$dem_params$spp1$lambda <- .1
Pop$dem_params$spp2$lambda <- .1
Pop$dem_params$spp3$lambda <- .1

#I PUT THIS BACK INSIDE RUN_SIM
# #Calculate movement probabilities (used to be in run_sim)
# Move_Prob  <- lapply(paste0("spp", seq_len(sim[["idx"]][["n.spp"]])), function(s) { move_prob_Lst(lambda = Pop[["dem_params"]][[s]][["lambda"]], hab = hab[["hab"]][[s]])})
# 
# 
# Move_Prob_spwn <- lapply(paste0("spp", seq_len(sim[["idx"]][["n.spp"]])), function(s) { move_prob_Lst(lambda = Pop[["dem_params"]][[s]][["lambda"]], hab = hab[["spwn_hab"]][[s]])})
# 
# 
# names(Move_Prob)      <- paste0("spp", seq_len(sim[["idx"]][["n.spp"]]))
# names(Move_Prob_spwn) <- paste0("spp", seq_len(sim[["idx"]][["n.spp"]]))

fields::image.plot(Pop$Start_pop[[1]], main = "spp1 starting biomass")
fields::image.plot(Pop$Start_pop[[2]], main = "spp1 starting biomass")
fields::image.plot(Pop$Start_pop[[3]], main = "spp1 starting biomass")

```

## Population movement

Now we set up the population tolerance to different temperatures which
determines how the populations move during the course of a year. We can then
plot the combined spatiotemporal suitable habitat to examine how these
interact.

```{r temp}
#load temp gradient

#moveCov <- readRDS(file="20 year moveCov matrices/GeorgesBank/GB_22yr_ConstTemp_HaddockStrata")
moveCov <- readRDS(file="20 year moveCov matrices/GeorgesBank/GB_22yr_ConstTemp_HaddockStrata_res2")

#set temperature preferences manually. 
#The following assumes moveCov has been created and already has an empty spp_tol sublist
moveCov[["spp_tol"]] <- list() #just in case
moveCov[["spp_tol"]] <- list("spp1" = list("mu" = 9, "va" = 4),  #Haddock
     "spp2" = list("mu" = 8.75, "va" = 4.25),  #Cod
     "spp3" = list("mu" = 9, "va" = 4) )    #Yellowtail

```

## Fleet models

Here we initialise the fleet with fish landings price per tonne, catchability
coefficients per population, fuel cost, the coefficients for the step function
and fleet behaviour.

We can plot the behaviour of the step function to check its suitable for our
simulations. This determines the relationship between the monetary value gained
from a fishing tow and the next move by the vessel when using the correlated
random walk function.

```{r fleets}

library(MixFishSim)

#no fishing
fleets <- init_fleet(sim_init = sim, VPT = list("spp1" = 0, "spp2" = 0, "spp3" = 0), #VPT = value per ton
                     Qs = list("fleet 1" = c("spp1" = 0, "spp2" = 0, "spp3" = 0)   #Q = catchability
                     ),
                     fuelC = list("fleet1" = 3),
                     step_params = list("fleet 1" = c("rate" = 3, "B1" = 1, "B2" = 2, "B3" = 3)
                     ),				
                     past_knowledge = FALSE,  #dont use past knowledge
                     past_year_month = TRUE,
                     past_trip = TRUE,
                     threshold = 0.7
)

```

## Run simulation

Finally we run the simulation. The output is a list of objects containing all
the information on fisheries catches, the population dynamics and population
distributions. These can be examined with some inbuilt plotting functions.

```{r sim}
run_simulation <- function(x){

  source("R/run_sim.R")

  #to source a new go_fish where I edited to skips most things:
  #1: load file
  source("R/go_fish_Bens.R") #my edited version that skips most things
  #2: allow the function to call other hidden functions from mixfishsim 
  environment(go_fish_Bens) <- asNamespace('MixFishSim')
  #3: replace go_fish with go_fish_Bens in the MixFishSim package
  assignInNamespace("go_fish", go_fish_Bens, ns = "MixFishSim")

#   #rebuild package
devtools::build()

  #load rcpp exports
Rcpp::sourceCpp(file= "src/Movement.cpp")
Rcpp::sourceCpp(file= "src/RcppExports.cpp")
Rcpp::compileAttributes() #this updates RcppExports.R file, which contains function definitions

  library(MixFishSim) #each core needs to load library

  #load data from CPU
  #  load("C:/Users/benjamin.levy/Desktop/Github/READ-PDB-blevy2-toy/20 year moveCov matrices/Final/Final_moveCov_12_9_Bensmethod.RData")

  #load constant temp data from Mars
  #load("/net/home5/blevy/Bens_R_Projects/READ-PDB-blevy2-toy/20 year moveCov matrices/Final/Final_constanttemp_20yr.RData")

  #load increase temp data from Mars
  load("/net/home5/blevy/Bens_R_Projects/READ-PDB-blevy2-toy/20 year moveCov matrices/Final/Increase_temp_20yr_LargerTempVA7.RData")

  #this starts the profiling
library(profvis)

profvis({

start_time <- Sys.time() # record start time

  res<- run_sim(sim_init = sim,
                pop_init = Pop,
                move_cov = moveCov,
                fleets_init = fleets,
                hab_init = hab,
                save_pop_bio = TRUE,
                survey = NULL, #surv_random, will try to survey after simulation
                closure = NULL,
                InParallel = TRUE,
                nz = nonzero_idx)  #does it runin parallel? Doesnt seem like it

end_time <- Sys.time() # record end time
tot_time <- end_time - start_time # calculate duration of lrren() example

})

}

#run in parallel

library(parallel)

nCoresToUse <- detectCores() - 1   #this show 16 cores but I think I have 6??

nCoresToUse <- 6

cl <- parallel::makeCluster(nCoresToUse,revtunnel = TRUE, outfile = "", verbose = TRUE, master=nsl(Sys.info()['nodename'])) #options from https://stackoverflow.com/questions/41639929/r-cannot-makecluster-multinode-due-to-cannot-open-the-connection-error

result <- list()

result <- parallel::parLapply(cl,1:6,run_simulation)
parallel::stopCluster(cl)

# 
# #below throws an error
# #result <- foreach(i=1:3) %dopar% run_simulation(i)
# 
# #below throws a similar error: 3 nodes produced errors; first error: missing value where TRUE/FALSE needed
#  parLapply(cl,1:3,run_simulation)
# stopCluster(cl)

```

## Summary plots

There are a series of input plotting functions to visualise the results of the
simulation. For example, we can explore:

  * the population dynamics for each species
* Seasonal patterns in exploitation
* the location choice of a vessel
* the realised step function for a vessel

Users will wish to define their own plots, depending on the issues of interest
and all the results are saved in the output from the run_sim function.

```{r plots}

## Biological
source("R/plot_pop_summary.R")
p1 <- plot_pop_summary(results = res,
                       timestep = "annual",
                       save = FALSE, 
                       save.location = NULL
)

plot_pop_summary(results = res, timestep = "daily", save = FALSE, save.location = "C:/Users/benjamin.levy/Desktop/Github/READ-PDB-blevy2-toy/testfolder")
p1

p2 <- plot_daily_fdyn(res)
p2

## Fishery

logs <- combine_logs(res[["fleets_catches"]])

p3 <- plot_vessel_move(sim_init = sim, logs = logs, fleet_no = 1, vessel_no = 5,
                       year_trip = 5, trip_no = 10)
p3

p4 <- plot_realised_stepF(logs = logs, fleet_no = 1, vessel_no = 1)
p4

#try some new ones

#plot survey 2 ways
p5 <- plot_survey(survey = res$survey, type = "spatial")
p5

p6 <- plot_survey(survey = res$survey, type = "index")
p6

#spatio temporal population plots
source("R/Bens_plot_pop_spatiotemp.R")
p7 <- Bens_plot_pop_spatiotemp(results = res,
                               save = TRUE, 
                               save.location = "testfolder/spatialplots"
)
p7

```

Note in our example how the fishing mortality rate for species 2 changes
following the spatial closure, which was set to cover some of the core
distribution of the population.