### 2023-01-03 Manuscript/Draft1cml.Rmd
example of working references

---
csl: canadian-journal-of-fisheries-and-aquatic-sciences.csl
output:
  bookdown::pdf_document2:
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes
  pdf_document:
    toc: no
  word_document: default
  html_document: default
bibliography: test.bib
fontsize: 12pt
header-includes:
- \usepackage{setspace}
- \doublespacing
- \usepackage{lineno}
- \linenumbers
- \usepackage[belowskip=0pt,aboveskip=0pt]{caption}  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
if(!require(kableExtra)) {  
  install.packages("kableExtra")
  require(kableExtra)}

#read in results tables
YT_results<-read.csv("Yellowtail_results.csv")
Cod_results<-read.csv("Cod_results.csv")
Had_results<-read.csv("Haddock_results.csv")
```

\newpage

# Estimating Population Trends with Stratified Random Sampling Under the Pressures of Climate Change

Benjamin A. Levy^1^, Christopher M. Legault^2^, Timothy J. Miller^2^, Elizabeth N. Brooks^2^

^1^Ben's Institution, USA  
^2^National Marine Fisheries Service, Northeast Fisheries Science Center, Woods Hole, MA, USA

Corresponding author: Ben Levy (benjamin.levy@noaa.gov)

Competing interests: The authors declare there are no competing interests.

<!-- Table1 <- YT_results -->

<!-- write.table( -->
<!--    format(f, digits=2), "", -->
<!--    sep="|", row.names=F, col.names=F, quote=F, eol="|\n") -->

\newpage

## Abstract

An Abstract

## Keywords

keyword 1, keyword 2

\newpage

\section{Introduction}

\section{Methods}

-- Used MixFishSim. Describe edits made to package

We use the R package *MixFishSim* (MFS) to model our populations [@dolder2020highly]. MFS is a discrete spatiotemporal simulation tool where users can model multiple species under varying environmental conditions. The package uses a delay-difference population model with discrete processes for growth, death, and recruitment of the population. We formulate the following inputs for the MFS package to address our research question.

## References

<div id="refs"></div>

\pagebreak

## Tables
</font>